EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-05-15 draft.tex
More comments addressed
Moving question about cloud to reproducible
@JJ changed the name of the section
Deleting resolved comments
Updating the now Reproducible section
I'm keeping just two roles. I will not mention installation.

  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
  % I like cloud-native more than event-based. 
  % How about that one?  - M
  % better than the previous one.

@@

\begin{abstract} 
Splitting a population into multiple instances is a technique that has
been used extensively in recent years in order to help improve the
performance of nature-inspired optimization algorithms. Work on those
populations can be done in parallel, and they can interact asynchronously,
a fact that can be leveraged to create scalable implementations based
on, among other methods, distributed, multi-threaded, parallel, and
cloud-native computing.  However, the design of these cloud-native,
distributed, multi-population algorithms is not a trivial task. Coming
from monolithic (single-instance) solutions, adaptations at several
levels, from the algorithmic to the functional, must be made to
leverage the scalability, elasticity, (limited) fault-tolerance,
reproducibility, and cost-effectiveness of cloud systems while, at the
same time, conserving the intended functionality. Instead of an
evolutive approach, in this paper we propose to create from scratch a
cloud-native optimization framework that is able to include multiple
algorithms and use few parameters to work. This solution goes beyond
current state of the art via a framework that is able to support
different algorithms at the same time, work asynchronously, and also
be easily deployable to any cloud platform. We evaluate the
performance and scalability of this solution, together with the effect
other design parameters had on it; in particular, the number and the
size of populations with respect to problem size. The architecture and
the implemented platform is an excellent alternative for running
locally or in the cloud, thus proving that cloud-native bioinspired
algorithms perform better in their ``natural'' environment than other
kinds of algorithms, % Did we compare it to other kind of algorithms?
                     % If we didn't, this should go - JJ 
                     % Compared against GA and PSO and others from BBOB - Mario
                     % on the algorithm, not speedup 
and setting a new baseline for scaling and
performance in the cloud.
\end{abstract}

@@

% But first, we give a brief description of the technology we used.

% Add if needed
%\subsection{Containers} 
%\label{containers}
% TO DO

%\subsection{Docker Compose} 
%\label{compose}
% TO DO

% Do we need a description of every container? Wouldn't a general 
% description  of the containers used be enough? - JJ 
% Ok, I think only workers and the controller need to be described in detail - Mario 
\subsection{Message Queue Container} 
\label{message_container}

@@

\section{Experimental Workflow in EvoSwarm} % We could upgrade this to a section and explain in detail 
% taking care of #5 - M
\label{experiment_flow}

% You can't start with this. Explain what this section is about, and link it with the previous one. Maybe it's better to create a big section with "Experimental setup", and make this a subsection, with subsections becoming subsubsections - JJ
Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
in Evo\-Swarm, and the three roles users can play. The first step is to develop the
search strategy algorithms that are going to be used by the multi-population
algorithm; algorithm developers do this. Then administrators are responsible for
the installation and configuration of the application. Once the application is
running, users can start one or several experiments, monitor the execution, and
analyze the results. Depending on the type of experiment executed, users could
have additional needs, in this case there are additional scripts for
processing the results of experiments to be used with the COCO framework. In the
next sections, we give more details about these roles and their workflows.

@@

For the secondary level of integration, developers need to add code to the setup

container
in the current configuration message (see Listing \ref{code:config})
workers executing a GA, initialize the mutation and crossover 
probabilities randomly from a specific range of values. Users specify
this range on the experiment configuration. 
Migration between populations is independent of the type of algorithm, as the
controller treats messages as the same type of objects. But developers could
also change the controller to specify new rules of operation, for instance, to
change the type of migration depending on the algorithm. 
For example, in the case of swarm-based algorithms,  migration between  
populations (swarms) can follow a topology, indicating valid connections 
between them (i.e., ring or hypercube). The controller
can apply these rules selectively only to swarm-based algorithms. 
In general, these configuration options will be algorithm and library specific.

@@

% You said "subsections will talk about roles", this is not a role. Change the title to that of a role. Would that be administrator? - JJ

\subsection{EvoSwarm Configuration}
\label{sec:evoswarm:config}
The administrator role is responsible for
editing the Docker Compose configuration files that reference the images created
by developers. To deploy the application, he or she needs to edit the {\tt
docker-compose.yml} file to specify the type and number of containers required
for the experiments. The {\tt docker-compose} application is responsible for
starting all the containers, and downloading new versions if it is necessary,
by executing the command {\tt docker-compose up}  inside the EvoSwarm directory

@@

% Again, not a role - JJ
\subsection{Experiment Execution}
\label{sec:exeperimentalDefinition}

@@

% This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
% This is one of the applications described in the Workflow Figure. But this is better explained in other places.
% We can delete it (even from the Figure)  
% Will leave it until we decide - M 
\subsection{Using the COCO Framework}

@@

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
\end{lstlisting}
% Do you need the details about the subnets? As long as they are here,
% you'll have to explain them. Why two subnets? It's probably better to
% leave that network configuration out anyway - JJ
% Ok, I removed that section. those parameters are required, but are used 
% internally by the cluster. - M 
% The security group was required to have external access to redis. - M
% Please clarify that those images are published in Docker hub by yourself - JJ
% I have done this a few commits back. (4 paragraphs up) - M

@@

    controller:
      image: mariosky/controller
      command: python controller.py
      environment:
          PYTHONUNBUFFERED: 1
          REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: controller
  \end{lstlisting}
% You will probably need to explain this a bit - JJ 
% done - M  
\subsubsection{Experimental Results}

@@

% JJ these are the three experiments, please check them. The last two addressing
% issues from a reviewer - M

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service? % This does not fit in
                                % here... It's related to
                                % reproducibility, and as such belongs
                                % in the previous section. The
                                % subsections that answer this should
                                % also go to the previous section - JJ
% Still need to check this out - JJ
\end{itemize}

% Also, point forward to the section (or subsections) that will be devoted to proving each point. BTW, they should probably be subsections.

@@

The experiments can be easily reproduced by using the configuration
files the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
             % Done, a few lines down - M
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config}. 
Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We mention earlier that we replicated the params of another paper - M

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ

@@

%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalspersecond}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }

  % Which function? - JJ
  \label{fig:spworker}
\end{figure}

@@

% No idea where this fits in. I think it's the second question in 7, so it should probably be a subsection - JJ
\subsection{Performance of Heterogeneous Populations in the BBOB Benchmark}
\label{sec:exp2}

@@

% If this is what you want to prove, it's not a very strong point. Maybe it
% would be stronger if you proved that these mixed algorithms can be tested in the
% same conditions as single-algorithm, and also that these mixed algorithms work
% better than single-algorithm versions - JJ
% Ok, stated like that - M

>>>>

  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}

@@

\begin{abstract} 
Splitting a population into multiple instances is a technique that has
been used extensively in recent years in order to help improve the
performance of nature-inspired optimization algorithms. Work on those
populations can be done in parallel, and they can interact asynchronously,
a fact that can be leveraged to create scalable implementations based
on, among other methods, distributed, multi-threaded, parallel, and
cloud-native computing.  However, the design of these cloud-native,
distributed, multi-population algorithms is not a trivial task. Coming
from monolithic (single-instance) solutions, adaptations at several
levels, from the algorithmic to the functional, must be made to
leverage the scalability, elasticity, (limited) fault-tolerance,
reproducibility, and cost-effectiveness of cloud systems while, at the
same time, conserving the intended functionality. Instead of an
evolutive approach, in this paper we propose to create from scratch a
cloud-native optimization framework that is able to include multiple
algorithms and use few parameters to work. This solution goes beyond
current state of the art via a framework that is able to support
different algorithms at the same time, work asynchronously, and also
be easily deployable to any cloud platform. We evaluate the
performance and scalability of this solution, together with the effect
other design parameters had on it; in particular, the number and the
size of populations with respect to problem size. The architecture and
the implemented platform is an excellent alternative for running
locally or in the cloud, thus proving that cloud-native bioinspired
algorithms perform better in their ``natural'' environment than other
kinds of algorithms, % Did we compare it to other kind of algorithms?
                     % If we didn't, this should go - JJ 
                     % Compared against GA and PSO and others from BBOB - Mario
                     % on the algorithm, not speedup 
                     % I like the sentence, If we did not prove it, 
                     % at least we build a tool that will allow
                     % algorithms to run in their natural environment.
                     % The focus of the Journal after all is in computer systems - M

and setting a new baseline for scaling and
performance in the cloud.
\end{abstract}

@@

\subsection{Message Queue Container} 
\label{message_container}

@@

\section{Workflow for Reproducible Experiment Design in EvoSwarm} 
\label{experiment_flow}
% Changed the section title to give more focus on reproducibility - M
% You can't start with this. Explain what this section is about, and link it with 
% the previous one. Maybe it's better to create a big section with "Experimental setup", 
% and make this a subsection, with subsections becoming subsubsections - JJ 
% done - M

An essential goal of our EvoSwarm application is to allow researchers the
description and orchestration of reproducible multi-population experiments.
Following the application design described in the previous section, we now
present the workflow of an experiment's entire life-cycle viewed from the
perspective of the two roles users can play. On one side is the user role that
designs and runs experiments and, on the other, the role of a developer of
algorithms. The scenario presented in Figure~\ref{fig:experiment_flow} shows the
workflow of an experiment in Evo\-Swarm the two roles users can play. The first
step is to develop the search strategy algorithms that are going to be used by
the multi-population algorithm; algorithm developers do this. Then users are
responsible for the configuration and deployment of the resources needed for the
experiment. Once the application is running, users can start one or several
experiments, monitor the execution, and analyze the results. Depending on the
type of experiment executed, users could have additional needs, in this case
there are additional scripts for processing the results of experiments to be
used with the COCO framework. In the next sections, we give more details about
these roles and their workflows, and at the end we present a proof-of-concept example of a
reproducible experiment in EvoSwarm.

%\begin{itemize}
  % Question move to here for the moment, I mention this in this intro. - Mario

%  \item Can EvoSwarm as a cloud-native, container-based application,  be deployed
%  locally or in the cloud by just specifying the configuration of container images
%  and resources to an orchestration service? 
%\end{itemize}

@@

For the secondary level of integration, developers need to add code to the setup
container to specify additional configuration options. For instance,
in the current configuration message (see Listing \ref{code:config})
workers executing a GA, initialize the mutation and crossover 
probabilities randomly from a specific range of values. Users specify
this range on the experiment configuration. 
Migration between populations is independent of the type of algorithm, as the
controller treats messages as the same type of objects. But developers could
also change the controller to specify new rules of operation, for instance, to
change the type of migration depending on the algorithm. 
For example, in the case of swarm-based algorithms,  migration between  
populations (swarms) can follow a topology, indicating valid connections 
between them (i.e., ring or hypercube). The controller
can apply these rules selectively only to swarm-based algorithms. 
In general, these configuration options will be algorithm and library specific.

@@

% You said "subsections will talk about roles", this is not a role. 
% Change the title to that of a role. Would that be administrator? - JJ

\subsection{EvoSwarm from the User Point of View}

\subsubsection{EvoSwarm Configuration}
\label{sec:evoswarm:config}
The administrator role is responsible for
editing the Docker Compose configuration files that reference the images created
by developers. To deploy the application, he or she needs to edit the {\tt
docker-compose.yml} file to specify the type and number of containers required
for the experiments. The {\tt docker-compose} application is responsible for
starting all the containers, and downloading new versions if it is necessary,
by executing the command {\tt docker-compose up}  inside the EvoSwarm directory

@@

% Again, not a role - JJ
\subsubsection{Experiment Execution}
\label{sec:exeperimentalDefinition}

@@

% This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
% This is one of the applications described in the Workflow Figure. But this is better explained in other places.
% We can delete it (even from the Figure)  
% Will leave it until we decide - M 
\subsubsection{Using the COCO Framework}

@@

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
\end{lstlisting}

@@

    controller:
      image: mariosky/controller
      command: python controller.py
      environment:
          PYTHONUNBUFFERED: 1
          REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: controller
  \end{lstlisting}

@@

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increase the performance of the
search by needing fewer function evaluations than a homogeneous setting?

% Moved reproducibility question from here - M
\end{itemize}

% Also, point forward to the section (or subsections) that will be devoted to proving each point.
% BTW, they should probably be subsections.

@@

The experiments can be easily reproduced by using the configuration
files the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. 
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config}. 
Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm.

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2.

@@

%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalspersecond}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
  \label{fig:spworker}
\end{figure}

@@

% No idea where this fits in. I think it's the second question in 7,
% so it should probably be a subsection - JJ

\subsection{Performance of Heterogeneous Populations in the BBOB Benchmark}
\label{sec:exp2}

@@
